---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Bivariate Analysis 
```{r bivariate_data}
load("../data/preprocessed/train.Rdata")
target <- 'price.log'
```

### Continuous Variables
```{r bi_quant_predictors}
# Obtain numeric variables
vars <- colnames(df)[sapply(df, is.numeric)]
vars <- vars[vars != target]

```

#### Target Correlation Analysis
```{r target_cor}
analysis <- corAnalysis(df, target = target)
analysis <- analysis %>% filter(abs(r) > .5 & (aR2.x > .5 | aR2.y > .5)) %>% arrange(-abs(r))
knitr::kable(analysis, digits = 4) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```

#### Predictor Regression Summary
```{r prs_quant, results='asis'}
analysis <- analyzR::multiSLM(df, vars, target = target)
models <- analysis$models
prsQuant <- analysis$summary
knitr::kable(prsQuant, digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```

#### Predictor Regression Analysis
```{r bi_quant_regression, fig.height=10, results='asis'}
for (i in 1:length(models)) {
  # Print regression results and statistics
  print(knitr::kable(models[[i]]$modelFit, digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right"))
  print(knitr::kable(models[[i]]$modelStats, digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right"))
  
  # Print regression diagnostic plots
  par(mfrow=c(3,2))
  plot(df[[vars[i]]], df[[target]], main = paste(target, "vs", vars[i]), xlab = vars[i], ylab = target)
  plot(models[[i]]$model$fitted.values, models[[i]]$model$model$price.log, main = "Predicted vs True", xlab = 'Predicted Log Price', ylab = "True Log Price")
  plot(models[[i]]$model)
}
```

#### Predictor Correlation Analysis
```{r corr}
corrData <- analyzR::correlation(df, target = "price.log", threshold = 0.5)
# Plot correlation matrix
corrplot(corrData$matrix, diag = FALSE,  tl.pos = "td", tl.cex = 0.5, method = "color", type = "upper")

knitr::kable(corrData$table, digits = 4) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```

### Categorical Variables
```{r bi_qual_predictors}
# Obtain numeric variables
vars <- colnames(df)[sapply(df, is.factor)]
```

#### Predictor Regression Summary
```{r prs_qual, results='asis'}
analysis <- analyzR::multiSLM(df, vars, target = target)
models <- analysis$models
prsQual <- analysis$summary
knitr::kable(prsQual, digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```

#### Predictor Regression Analysis
```{r bi_qual_regression, fig.height=10, results='asis'}
for (i in 1:length(models)) {
  # Print regression results and statistics
  print(knitr::kable(models[[i]]$modelFit, digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right"))
  print(knitr::kable(models[[i]]$modelStats, digits = 3) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right"))
  
  # Print regression diagnostic plots
  print(plotFactorHist(df, target, vars[i], xLab = "Log Price", facets = TRUE))
  par(mfrow=c(3,2))
  plot(df[[vars[i]]], df[[target]], main = paste(target, "vs", vars[i]), xlab = vars[i], ylab = target)
  plot(models[[i]]$model$fitted.values, models[[i]]$model$model$price.log, main = "Predicted vs True", xlab = 'Predicted Log Price', ylab = "True Log Price")
  plot(models[[i]]$model)
}
```


#### Predictor Association Analysis
```{r target_as}
assoc <- analyzR::association(df, target = target, threshold = 0.5)

# Plot association matrix
corrplot(assoc$matrix, diag = FALSE,  tl.pos = "td", tl.cex = 0.5, method = "color", type = "upper")

# Print asociationtable
knitr::kable(assoc$table, digits = 4) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```

## Multivariate Analysis
```{r qual_neighborhood}
plotScatter(df, xVar = 'Overall.Qual', yVar = 'price.log', groupVar = 'Neighborhood',
            plotTitle = "Log Price by Overall Quality and Neighborhood")
```

### Summary
#### Relative Importance

```{r relative_importance}
ri <- rbind(prsQual, prsQuant)
ri <- ri %>% arrange(-adj.r.squared)
print(plotBar(ri[1:20,], xVar = 'term', xLab = "Term", yVar = 'adj.r.squared', yLab = "Proportion Variance Explained", values = FALSE, plotTitle = "Relative Importance of Top 20 Terms"))
```

#### Correlation of Features
```{r corr_top}
topFeatures <- ri$term[1:30] 
df2 <- df %>% dplyr::select(topFeatures)
corrData <- analyzR::correlation(df2, target = "price.log", threshold = 0.5, bigMatrix = 30)
# Plot correlation matrix
corrplot(corrData$matrix, diag = FALSE,  tl.pos = "td", tl.cex = 0.5, method = "color", type = "upper", tl.col = 'black')

knitr::kable(corrData$table, digits = 4) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```

#### Association of Top Features
```{r ass_top}
assoc <- analyzR::association(df2, target = target, threshold = 0.5, bigMatrix = 30)

# Plot association matrix
corrplot(assoc$matrix, diag = FALSE,  tl.pos = "td", tl.cex = 0.5, method = "color", type = "upper", tl.col = 'black')

# Print asociationtable
knitr::kable(assoc$table, digits = 4) %>%
  kableExtra::kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = T, position = "float_right")
```
